%!TEX root = ../paper.tex

% ---------------------------------------------------------------------------- %
% Outline

\iftoggle{outline}{

    \begin{itemize}
        \item VMs are good
        \item Hard to build, even with VM gramworks like Truffle \& pypy
        \item Makes it hard for a new langauge to get a VM
        \item We are exploring making a VM for Grace by adapting VM for NS
        \item initial results are positive
    \end{itemize}

}

\iftoggle{oldcomments}{
	\sm{I feel this characterization is not quite accurate. At least for RPython, each interpreter gets its own VM generated. With Truffle, we indeed reuse the JVM as a host VM.}
	\sm{The use of macros seems highly over-engineered...}
	\sm{and, you might like the xspace package, I think, to avoid always having to put braces at the end}
	\sm{really not a fan of macros. This makes editing really hard!}
}

% 
% ---------------------------------------------------------------------------- %



% ---------------------------------------------------------------------------- %
% 

With Truffle and RPython, language implementation frameworks have emerged that enable us to
implement languages with the performance of state-of-the-art \vm/s by building
simple interpreters.
% \vm/ frameworks have emerged and are quickly becoming a core component to language implementation. 
% Truffle and RPython are \VM/ frameworks: 
With Truffle, \AST/ interpreters are built on top of a host \VM/, which provides for instance garbage collection and just-in-time compilation.
% which a developer uses to author a simple \AST/ interpreter and through doing so automatically gain features such, and optimizations to the \AST/; and
With the RPython toolchain, a tailored \vm/ is generated based on a simple interpreter. The toolchain adds for instance garbage collection and a tracing
just-in-time compiler to enable competitive performance.
Unsurprisingly, developers have already used these frameworks to successfully implement popular object-oriented programming languages (such as Python, Ruby, or JavaScript), with performance results as good as state of the art \JITing/ \VMs/ \citep{Marr2015,Marr:2016:AWFY}.

Despite the multitude of language implementations built using these frameworks, it appears that developers tend to recreate similar components (e.g., basic control structures, \AST/ transformations, object storage) \emph{on-top-of} the framework rather than reuse existing implementations; perhaps doing so to gain fine-grained control over later optimizations (at the cost of a significant amount of work).

Recently, the notion of developing a set of components that can be easily extended to create a tailored \VM/s have gained traction. For example, \rcitea{WoB2014}\mwh{The citation style breaks this reference} introduced a robust \emph{Object Storage Model}, which has now been integrated into the Truffle framework; and \citet{Inostroza2015} propose reusable components for languages with denotational semantics using \emph{Object Algebras} \rcite{Oliveira2012}. From much the same perspective, we propose to explore the potential to create new language implementations by extending existing ones built using language implementation frameworks.
As an initial experiment, we adapt SOMns to realize a \VM/ for Grace.


%
% ---------------------------------------------------------------------------- %



% ---------------------------------------------------------------------------- %
% Notes

% Stefan - People are playing with reusable components built in VM frameworks
%
% eg Product Lines of Interpreters Using Truffle with Object Algebras
% 	 Modular interpreters for the masses: implicit context propagation using object algebras

% Stefan - Object storage is provided by truffle, so, the fact that SOMns has its own is somewhat an artifact of ‘history’, and perhaps somewhat an optimization/step to avoid dependence on closed-source Oracle features. There are other aspects like storage strategies (think of our arrays), which haven’t been included in truffle yet, but are for instance in RPython: 
% -- http://dl.acm.org/citation.cfm?id=2816716

% Richard - The original idea of a \emph{\VM/} was a tool that compiled a program from an abstract into machine executable code, useful in that developers could implement a method to parse their language into the \ir understood by the \VM/. 

    % Stefan - I don’t think so, p-code might be the first example of a virtual machine. There was no translation, it was just to avoid targeting a specific hardware machine, because back then they had many more different architectures than 3-4 people care about today


% Richard - An interpreter is collections of tools for executing a program. Lexical analysis encodes an intermeditary representation with the logic of the program and then that representation is visited to realize execution. From its outset, practitioners have created a multitude of different interpreters for different languages, each with a toolkit based on similar conventions but unique in its implementation (often required to support the semantics of the language being interpreted). 

% 
% ---------------------------------------------------------------------------- %
